---
title: "Data 621 Final"
author: "Matthew Lucich"
date: "11/21/2021"
output: html_document
---

```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r, include=FALSE}
library(tidyverse)
library(imputeMissings)
library(vars)
library(TSstudio)
library(tseries)
library(corrplot)
library(stargazer)
library(Metrics)
library(psych)
```


## Data Cleaning

```{r}

df_wb <- read_csv("world-bank.csv")

#glimpse(df_wb)

# Filter for Brazil, Russia, India, China, and South Africa 
df_brics <- df_wb %>% filter(grepl("BRA|CHN|IND|RUS|ZAF", `Country Code`))

variables_to_include <- "NY.GDP.PCAP.KD.ZG|FS.AST.CGOV.GD.ZS|FP.CPI.TOTL|NY.GDP.MKTP.KD.ZG|MS.MIL.XPND.GD.ZS|SE.PRM.AGES|DT.DOD.DSTC.ZS|SP.POP.TOTL|GB.XPD.RSDV.GD.ZS"
df_brics <- df_brics %>% filter(grepl(variables_to_include, `Series Code`))

#glimpse(df_brics)

```


```{r, echo=FALSE}

#unique(df_brics$`Series Name`)

#colnames(df_brics)

df_brics_wide <- df_brics %>% rename(country = "Country Name",
                    country_code = "Country Code",
                    var = "Series Name",
                    var_code = "Series Code")

df_brics_wide <- df_brics_wide %>% rename_with(~ gsub("\\[(.*?)\\]", "", .x))
df_brics_wide <- df_brics_wide %>% rename_with(~ gsub(" ", "", .x))

#glimpse(df_brics_wide)

df_brics_wide['country_var'] <- paste(df_brics_wide$country, df_brics_wide$var)

```




```{r}

df_brics_trim <- df_brics_wide %>% dplyr::select(-c(country, country_code, var, var_code))

#transpose(df_brics_trim)

# first remember the names
n <- df_brics_trim$country_var

# transpose all but the first column (name)
df_brics_t <- as.data.frame(t(df_brics_trim[,-1]))
colnames(df_brics_t) <- n

df_brics_t <- df_brics_t %>% rename_with(~ gsub("\\((.*?)\\)", "", .x))
df_brics_t <- df_brics_t %>% rename_with(~ gsub(" ", "_", .x))
df_brics_t <- df_brics_t %>% rename_with(~ gsub(",", "", .x))
df_brics_t <- df_brics_t %>% rename_with(~ gsub("\\.", "", .x))
df_brics_t <- df_brics_t %>% rename_with(~ gsub("\\_$", "", .x))
df_brics_t <- df_brics_t %>% rename_with(~ tolower(gsub("-", "_", .x)))
df_brics_t <- df_brics_t %>% rename_with(~ gsub("n_federation", "", .x))
df_brics_t <- df_brics_t %>% rename_with(~ gsub("_etc", "", .x))

#head(df_brics_t)

```


```{r}

df_brics_t[df_brics_t==".."] <- NA

# Visualize NA counts for each column
df_brics_t  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()


df_bics <- df_brics_t %>% dplyr::select(!starts_with("Russia"))

#head(df_bics, 10)

dim(df_bics)

df_bics <- df_bics[10:49, ]

#head(df_bics)

#write.csv(df_bics,'bics-raw.csv')

```


```{r, echo=FALSE}

# Visualize NA counts for each column
df_bics  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()

```


```{r, echo=FALSE}

df_bics_imp <- df_bics %>% impute(object = NULL, method = "median/mode", flag = FALSE)

head(df_bics_imp)

describe(df_bics_num)

```


```{r, echo=FALSE}

# Convert columns to numeric
df_bics_num <- as.data.frame(lapply(df_bics_imp, as.numeric))

# Histograms
df_bics_num %>%
  gather() %>%
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density(fill = "steelblue", alpha=0.9, color="steelblue") +
    geom_histogram(aes(y=..density..), alpha=0.5, fill = "lightblue", color="lightblue", position="identity")

# Drop starting age variable due to low variance
df_bics_num <- df_bics_num %>% dplyr::select(!ends_with("_age"))


```



```{r}

# Visualize correlation between variables
corrplot(cor(df_bics_num %>% dplyr::select(starts_with("Brazil"))), method="shade", shade.col=NA, tl.col="black", tl.srt=45)

```



```{r, echo=FALSE}

braz_gdp <- ts(df_bics_num$brazil_gdp_growth, start = c(1981), frequency = 1)
braz_claims <- ts(df_bics_num$brazil_claims_on_central_government, start = c(1981), frequency = 1)
braz_cpi <- ts(df_bics_num$brazil_consumer_price_index, start = c(1981), frequency = 1)
braz_milt <- ts(df_bics_num$brazil_military_expenditure, start = c(1981), frequency = 1)
braz_st_debt <- ts(df_bics_num$brazil_short_term_debt, start = c(1981), frequency = 1)
braz_pop <- ts(df_bics_num$brazil_population_total, start = c(1981), frequency = 1)

india_gdp <- ts(df_bics_num$india_gdp_growth, start = c(1981), frequency = 1)
india_claims <- ts(df_bics_num$india_claims_on_central_government, start = c(1981), frequency = 1)
india_cpi <- ts(df_bics_num$india_consumer_price_index, start = c(1981), frequency = 1)
india_milt <- ts(df_bics_num$india_military_expenditure, start = c(1981), frequency = 1)
india_st_debt <- ts(df_bics_num$india_short_term_debt, start = c(1981), frequency = 1)
india_pop <- ts(df_bics_num$india_population_total, start = c(1981), frequency = 1)

china_gdp <- ts(df_bics_num$china_gdp_growth, start = c(1981), frequency = 1)
china_claims <- ts(df_bics_num$china_claims_on_central_government, start = c(1981), frequency = 1)
china_cpi <- ts(df_bics_num$china_consumer_price_index, start = c(1981), frequency = 1)
china_milt <- ts(df_bics_num$china_military_expenditure, start = c(1981), frequency = 1)
china_st_debt <- ts(df_bics_num$china_short_term_debt, start = c(1981), frequency = 1)
china_pop <- ts(df_bics_num$china_population_total, start = c(1981), frequency = 1)

so_af_gdp <- ts(df_bics_num$south_africa_gdp_growth, start = c(1981), frequency = 1)
so_af_claims <- ts(df_bics_num$south_africa_claims_on_central_government, start = c(1981), frequency = 1)
so_af_cpi <- ts(df_bics_num$south_africa_consumer_price_index, start = c(1981), frequency = 1)
so_af_milt <- ts(df_bics_num$south_africa_military_expenditure, start = c(1981), frequency = 1)
so_af_st_debt <- ts(df_bics_num$south_africa_short_term_debt, start = c(1981), frequency = 1)
so_af_pop <- ts(df_bics_num$south_africa_population_total, start = c(1981), frequency = 1)

```



# Brazil


## Plot Series (overlayed)

```{r, echo=FALSE}

ts.plot(diff(braz_gdp), diff(braz_claims), diff(braz_cpi), diff(braz_milt), diff(braz_st_debt),
        gpars=list(xlab="year", ylab="Variables", col = 1:5, lty=c(1:5)))
legend("topleft", legend = c("GDP", "Claims", "CPI", "Miltary", "Debt"), col = 1:5, lty=c(1:5))

```


## Test for Stationary

```{r, echo=FALSE}

pp.test(braz_claims)
pp.test(braz_cpi)
pp.test(braz_milt)
pp.test(braz_st_debt)

pp.test(diff(braz_claims))
pp.test(diff(braz_cpi))
pp.test(diff(braz_milt))
pp.test(diff(braz_st_debt))

```


## Create vectors of time-series objects for modeling

```{r, echo=FALSE}

braz_vec <- cbind(diff(braz_gdp), diff(braz_claims), diff(braz_cpi), diff(braz_milt), diff(braz_st_debt))
colnames(braz_vec) <- cbind("braz_gdp_diff","braz_claims_diff","braz_cpi_diff","braz_milt_diff", "braz_st_debt_diff")

```


## Train, test split

```{r, echo=FALSE}

split_brazil <- ts_split(ts.obj = braz_vec, sample.out = 4)

train_brazil <- split_brazil$train
test_brazil <- split_brazil$test

length(train_brazil)
length(test_brazil)

```


# VAR - Vector AutoRegressive model


## Find optimal lag length

```{r, echo=FALSE}

lagselect <- VARselect(braz_vec, lag.max = 10, type = "const")
lagselect$selection

```


## Build VAR model

```{r, echo=FALSE}

braz_model <- VAR(braz_vec, p = 3, type = "const", season = NULL, exog = NULL) 
summary(braz_model)

```


## Stargazer Output

```{r, echo=FALSE}

# 1) GDP 2) Claims 3) CPI 4) Milt 5) Debt
stargazer(braz_model[["varresult"]], type="text")

```


## Check for Serial Correlation

```{r}

braz_serial <- serial.test(braz_model, type = "PT.asymptotic")
braz_serial

```


## Test for Heteroscedasticity

```{r}

braz_arch <- vars::arch.test(braz_model, lags.multi = 15, multivariate.only = TRUE)
braz_arch

```


## Test for Normality of Residuals

```{r}

braz_norm <- normality.test(braz_model, multivariate.only = TRUE)
braz_norm

```


## Check for Stability of model

```{r}

braz_stability <- stability(braz_model, type = "OLS-CUSUM")
plot(braz_stability)

```


## Check for Granger Causality

```{r}

granger_gdp <- causality(braz_model, cause = "braz_gdp_diff")
granger_gdp

granger_claims <- causality(braz_model, cause = "braz_claims_diff")
granger_claims

granger_cpi <- causality(braz_model, cause = "braz_cpi_diff")
granger_cpi

granger_milt <- causality(braz_model, cause = "braz_milt_diff")
granger_milt

granger_debt <- causality(braz_model, cause = "braz_st_debt_diff")
granger_debt

```



## Create Impulse Response Functions

```{r}


#braz_all_irf <- irf(braz_model, response = "braz_gdp_diff", n.ahead = 4, boot = TRUE, run = 300, ci = 0.9, ortho = FALSE)
#plot(braz_all_irf)


#braz_all_irf <- irf(braz_model, response = "braz_gdp_diff", n.ahead = 4, boot = FALSE, ci = 0.9, ortho = FALSE)
#plot(braz_all_irf)


# Black line outside of confidence interval means the response is significant!
braz_claims_irf <- irf(braz_model, impulse = "braz_claims_diff", response = "braz_gdp_diff", n.ahead = 4, boot = TRUE, run = 300, ci = 0.9, ortho = FALSE)
plot(braz_claims_irf)

braz_cpi_irf <- irf(braz_model, impulse = "braz_cpi_diff", response = "braz_gdp_diff", n.ahead = 4, boot = TRUE, run = 300, ci = 0.9, ortho = FALSE)
plot(braz_cpi_irf)

braz_milt_irf <- irf(braz_model, impulse = "braz_milt_diff", response = "braz_gdp_diff", n.ahead = 4, boot = TRUE, run = 300, ci = 0.9, ortho = FALSE)
plot(braz_milt_irf)

# Debt has delayed impact on GDP?
braz_debt_irf <- irf(braz_model, impulse = "braz_st_debt_diff", response = "braz_gdp_diff", n.ahead = 4, boot = TRUE, run = 300, ci = 0.9, ortho = FALSE)
plot(braz_debt_irf)

```



## Review Variance Decomposition

```{r}

vd1 <- fevd(braz_model, n.ahead = 4)

plot(vd1)

```



## Generate Predictions

```{r}

forecast <- predict(braz_model, n.ahead = 4, ci = 0.95)
fanchart(forecast, names = "braz_gdp_diff", colors = c("seagreen","lightgreen"), main = "Forecasts", xlab = "Year", ylab = "GDP Growth %")

```

# View Actuals vs Predicted

```{r}

index_pred <- append(index(train_brazil[,1]), 2017:2019)

all_actuals <- append(train_brazil[,1], test_brazil[1:3,1])

all_preds <- append(train_brazil[,1], forecast$fcst$braz_gdp_diff[,1])

all_actuals_dediff <- diffinv(all_actuals, xi = braz_gdp[1])

all_preds_dediff <- diffinv(all_preds, xi = braz_gdp[1])
all_preds_dediff <- all_preds_dediff[-length(all_preds_dediff)]

all_actuals_dediff
all_preds_dediff

length(index_pred)
length(all_actuals_dediff)
length(all_preds_dediff[-1])

plot(index_pred,
     all_preds_dediff[-1],
     type = "l",
     ylab="Actuals and Predicted",
     xlab="Year",
     col = "red")
lines(index_pred,
      all_actuals_dediff[-1],
      type = "l",
      col = "green")
legend("topleft",
c("Predicted","Actuals"),
fill=c("red","green"))

```



## VAR - Evaluation Metrics

```{r}

# Root Mean Squared Error
print("Root Mean Squared Error")
rmse(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Error
print("Mean Absolute Error")
mae(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Percent Error
print("Mean Absolute Percent Error")
mape(all_actuals_dediff[37:39], all_preds_dediff[37:39])

```







# SVAR - Structural Vector AutoRegressive model

## SVAR - Build SVAR model

```{r, echo=FALSE}

amat <- diag(5)

amat[2,1] <- NA
amat[3,1] <- NA
amat[4,1] <- NA
amat[5,1] <- NA
amat[3,2] <- NA
amat[4,2] <- NA
amat[5,2] <- NA
amat[4,3] <- NA
amat[5,3] <- NA
amat[5,4] <- NA

braz_svar_model <- SVAR(braz_model, Amat = amat, Bmat = NULL, hessian = TRUE, estmethod = c("scoring", "direct"))
summary(braz_svar_model)

```


## SVAR - Review Variance Decomposition

```{r}

vd1 <- fevd(braz_svar_model, n.ahead = 4)

plot(vd1)

```


## SVAR - Generate predictions


```{r}

irfvc <- irf(braz_svar_model, n.ahead = 3, cumulative=TRUE)
summary(irfvc)

```


# SVAR - View Actuals vs Predicted

```{r}

index_pred <- append(index(train_brazil[,1]), 2017:2019)

all_actuals_svar <- append(train_brazil[,1], test_brazil[1:3,1])

all_preds_svar <- append(train_brazil[,1], irfvc$irf$braz_gdp_diff[,1])

all_actuals_dediff_svar <- diffinv(all_actuals_svar, xi = braz_gdp[1])

all_preds_dediff_svar <- diffinv(all_preds_svar, xi = braz_gdp[1])
all_preds_dediff_svar <- all_preds_dediff_svar[-length(all_preds_svar)]

plot(index_pred,
     all_preds_dediff_svar[-1],
     type = "l",
     ylab="Actuals and Predicted",
     xlab="Year",
     col = "red")
lines(index_pred,
      all_actuals_dediff_svar[-1],
      type = "l",
      col = "green")
legend("topleft",
c("Predicted","Actuals"),
fill=c("red","green"))

```

## SVAR - Evaluation Metrics

```{r}

# Root Mean Squared Error
print("Root Mean Squared Error")
rmse(all_actuals_dediff_svar[37:39], all_preds_dediff_svar[37:39])


# Mean Absolute Error
print("Mean Absolute Error")
mae(all_actuals_dediff_svar[37:39], all_preds_dediff_svar[37:39])


# Mean Absolute Percent Error
print("Mean Absolute Percent Error")
mape(all_actuals_dediff_svar[37:39], all_preds_dediff_svar[37:39])

```






# India


## Plot Series (overlayed)

```{r, echo=FALSE}

ts.plot(diff(india_gdp), diff(india_claims), diff(india_cpi), diff(india_milt), diff(india_st_debt),
        gpars=list(xlab="year", ylab="Variables", col = 1:5, lty=c(1:5)))
legend("topleft", legend = c("GDP", "Claims", "CPI", "Miltary", "Debt"), col = 1:5, lty=c(1:5))

```


## Test for Stationary

```{r, echo=FALSE}

pp.test(india_claims)
pp.test(india_cpi)
pp.test(india_milt)
pp.test(india_st_debt)

pp.test(diff(india_claims))
pp.test(diff(india_cpi))
pp.test(diff(india_milt))
pp.test(diff(india_st_debt))

```


## Create vectors of time-series objects for modeling

```{r, echo=FALSE}

india_vec <- cbind(diff(india_gdp), diff(india_claims), diff(india_cpi), diff(india_milt), diff(india_st_debt))
colnames(india_vec) <- cbind("india_gdp_diff","india_claims_diff","india_cpi_diff","india_milt_diff", "india_st_debt_diff")

```


## Train, test split

```{r, echo=FALSE}

split_india <- ts_split(ts.obj = india_vec, sample.out = 4)

train_india <- split_india$train
test_india <- split_india$test

length(train_india)
length(test_india)

```


# VAR - Vector AutoRegressive model


## Find optimal lag length

```{r, echo=FALSE}

lagselect <- VARselect(india_vec, lag.max = 10, type = "const")
lagselect$selection

```


## Build VAR model

```{r, echo=FALSE}

india_model <- VAR(india_vec, p = 4, type = "const", season = NULL, exog = NULL) 
summary(india_model)

```


## Stargazer Output

```{r, echo=FALSE}

stargazer(india_model[["varresult"]], type="text")

```


## Check for Serial Correlation

```{r}

india_serial <- serial.test(india_model, type = "PT.asymptotic")
india_serial

```


## Test for Heteroscedasticity

```{r}

india_arch <- vars::arch.test(india_model, lags.multi = 15, multivariate.only = TRUE)
india_arch

```


## Test for Normality of Residuals

```{r}

india_norm <- normality.test(india_model, multivariate.only = TRUE)
india_norm

```


## Check for Stability of model

```{r}

india_stability <- stability(india_model, type = "OLS-CUSUM")
plot(india_stability)

```


## Check for Granger Causality

```{r}

granger_gdp <- causality(india_model, cause = "india_gdp_diff")
granger_gdp

granger_claims <- causality(india_model, cause = "india_claims_diff")
granger_claims

granger_cpi <- causality(india_model, cause = "india_cpi_diff")
granger_cpi

granger_milt <- causality(india_model, cause = "india_milt_diff")
granger_milt

granger_debt <- causality(india_model, cause = "india_st_debt_diff")
granger_debt

```



## Create Impulse Response Functions

```{r}

# Black line outside of confidence interval means the response is significant!
india_claims_irf <- irf(india_model, impulse = "india_claims_diff", response = "india_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(india_claims_irf)

india_cpi_irf <- irf(india_model, impulse = "india_cpi_diff", response = "india_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(india_cpi_irf)

india_milt_irf <- irf(india_model, impulse = "india_milt_diff", response = "india_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(india_milt_irf)

# Debt has delayed impact on GDP?
india_debt_irf <- irf(india_model, impulse = "india_st_debt_diff", response = "india_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(india_debt_irf)


```



## Review Variance Decomposition

```{r}

vd1 <- fevd(india_model, n.ahead = 4)

plot(vd1)

```



## Generate Predictions

```{r}

forecast <- predict(india_model, n.ahead = 4, ci = 0.95)
fanchart(forecast, names = "india_gdp_diff", colors = c("seagreen","lightgreen"), main = "Forecasts", xlab = "Year", ylab = "GDP Growth %")

```

# View Actuals vs Predicted

```{r}

index_pred <- append(index(train_india[,1]), 2017:2019)

all_actuals <- append(train_india[,1], test_india[1:3,1])

all_preds <- append(train_india[,1], forecast$fcst$india_gdp_diff[,1])

all_actuals_dediff <- diffinv(all_actuals, xi = india_gdp[1])

all_preds_dediff <- diffinv(all_preds, xi = india_gdp[1])
all_preds_dediff <- all_preds_dediff[-length(all_preds_dediff)]

all_actuals_dediff
all_preds_dediff

length(index_pred)
length(all_actuals_dediff)
length(all_preds_dediff[-1])

plot(index_pred,
     all_preds_dediff[-1],
     type = "l",
     ylab="Actuals and Predicted",
     xlab="Year",
     col = "red")
lines(index_pred,
      all_actuals_dediff[-1],
      type = "l",
      col = "green")
legend("topleft",
c("Predicted","Actuals"),
fill=c("red","green"))

```

## VAR - Evaluation Metrics

```{r}

# Root Mean Squared Error
print("Root Mean Squared Error")
rmse(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Error
print("Mean Absolute Error")
mae(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Percent Error
print("Mean Absolute Percent Error")
mape(all_actuals_dediff[37:39], all_preds_dediff[37:39])

```













# China


## Plot Series (overlayed)

```{r, echo=FALSE}

ts.plot(diff(china_gdp), diff(china_claims), diff(china_cpi), diff(china_milt), diff(china_st_debt),
        gpars=list(xlab="year", ylab="Variables", col = 1:5, lty=c(1:5)))
legend("topleft", legend = c("GDP", "Claims", "CPI", "Miltary", "Debt"), col = 1:5, lty=c(1:5))

```


## Test for Stationary

```{r, echo=FALSE}

pp.test(china_claims)
pp.test(china_cpi)
pp.test(china_milt)
pp.test(china_st_debt)

pp.test(diff(china_claims))
pp.test(diff(china_cpi))
pp.test(diff(china_milt))
pp.test(diff(china_st_debt))

```


## Create vectors of time-series objects for modeling

```{r, echo=FALSE}

china_vec <- cbind(diff(china_gdp), diff(china_claims), diff(china_cpi), diff(china_milt), diff(china_st_debt))
colnames(china_vec) <- cbind("china_gdp_diff","china_claims_diff","china_cpi_diff","china_milt_diff", "china_st_debt_diff")

```


## Train, test split

```{r, echo=FALSE}

split_china <- ts_split(ts.obj = china_vec, sample.out = 4)

train_china <- split_china$train
test_china <- split_china$test

length(train_china)
length(test_china)

```


# VAR - Vector AutoRegressive model


## Find optimal lag length

```{r, echo=FALSE}

lagselect <- VARselect(china_vec, lag.max = 10, type = "const")
lagselect$selection

```


## Build VAR model

```{r, echo=FALSE}

china_model <- VAR(china_vec, p = 4, type = "const", season = NULL, exog = NULL) 
summary(china_model)

```


## Stargazer Output

```{r, echo=FALSE}

stargazer(china_model[["varresult"]], type="text")

```


## Check for Serial Correlation

```{r}

china_serial <- serial.test(china_model, type = "PT.asymptotic")
china_serial

```


## Test for Heteroscedasticity

```{r}

china_arch <- vars::arch.test(china_model, lags.multi = 15, multivariate.only = TRUE)
china_arch

```


## Test for Normality of Residuals

```{r}

china_norm <- normality.test(china_model)
china_norm

```


## Check for Stability of model

```{r}

china_stability <- stability(china_model, type = "OLS-CUSUM")
plot(china_stability)

```


## Check for Granger Causality

```{r}

granger_gdp <- causality(china_model, cause = "china_gdp_diff")
granger_gdp

granger_claims <- causality(china_model, cause = "china_claims_diff")
granger_claims

granger_cpi <- causality(china_model, cause = "china_cpi_diff")
granger_cpi

granger_milt <- causality(china_model, cause = "china_milt_diff")
granger_milt

granger_debt <- causality(china_model, cause = "china_st_debt_diff")
granger_debt

```



## Create Impulse Response Functions

```{r}

# Black line outside of confidence interval means the response is significant!
china_claims_irf <- irf(china_model, impulse = "china_claims_diff", response = "china_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(china_claims_irf)

china_cpi_irf <- irf(china_model, impulse = "china_cpi_diff", response = "china_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(china_cpi_irf)

china_milt_irf <- irf(china_model, impulse = "china_milt_diff", response = "china_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(china_milt_irf)

# Debt has delayed impact on GDP?
china_debt_irf <- irf(china_model, impulse = "china_st_debt_diff", response = "china_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(china_debt_irf)


```



## Review Variance Decomposition

```{r}

vd1 <- fevd(china_model, n.ahead = 4)

plot(vd1)

```



## Generate Predictions (differenced data)

```{r}

forecast <- predict(china_model, n.ahead = 4, ci = 0.95)
fanchart(forecast, names = "china_gdp_diff", colors = c("seagreen","lightgreen"), main = "Forecasts", xlab = "Year", ylab = "GDP Growth %")

```

# View Actuals vs Predicted

```{r}

index_pred <- append(index(train_china[,1]), 2017:2019)

all_actuals <- append(train_china[,1], test_china[1:3,1])

all_preds <- append(train_china[,1], forecast$fcst$china_gdp_diff[,1])

all_actuals_dediff <- diffinv(all_actuals, xi = china_gdp[1])

all_preds_dediff <- diffinv(all_preds, xi = china_gdp[1])
all_preds_dediff <- all_preds_dediff[-length(all_preds_dediff)]

all_actuals_dediff
all_preds_dediff

plot(index_pred,
     all_preds_dediff[-1],
     type = "l",
     ylab="Actuals and Predicted",
     xlab="Year",
     col = "red")
lines(index_pred,
      all_actuals_dediff[-1],
      type = "l",
      col = "green")
legend("topleft",
c("Predicted","Actuals"),
fill=c("red","green"))

```



## VAR - Evaluation Metrics

```{r}

# Root Mean Squared Error
print("Root Mean Squared Error")
rmse(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Error
print("Mean Absolute Error")
mae(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Percent Error
print("Mean Absolute Percent Error")
mape(all_actuals_dediff[37:39], all_preds_dediff[37:39])

```



# South Africa


## Plot Series (overlayed)

```{r, echo=FALSE}

ts.plot(diff(so_af_gdp), diff(so_af_claims), diff(so_af_cpi), diff(so_af_milt), diff(so_af_st_debt),
        gpars=list(xlab="year", ylab="Variables", col = 1:5, lty=c(1:5)))
legend("topleft", legend = c("GDP", "Claims", "CPI", "Miltary", "Debt"), col = 1:5, lty=c(1:5))

```


## Test for Stationary

```{r, echo=FALSE}

pp.test(so_af_claims)
pp.test(so_af_cpi)
pp.test(so_af_milt)
pp.test(so_af_st_debt)

pp.test(diff(so_af_claims))
pp.test(diff(so_af_cpi))
pp.test(diff(so_af_milt))
pp.test(diff(so_af_st_debt))

```


## Create vectors of time-series objects for modeling

```{r, echo=FALSE}

#so_af_vec <- cbind(diff(so_af_gdp), diff(so_af_claims), diff(so_af_cpi), diff(so_af_milt), diff(so_af_st_debt))
#colnames(so_af_vec) <- cbind("so_af_gdp_diff","so_af_claims_diff","so_af_cpi_diff","so_af_milt_diff", "so_af_st_debt_diff")

so_af_vec <- cbind(diff(so_af_gdp), diff(so_af_claims), diff(so_af_cpi), diff(so_af_milt))
colnames(so_af_vec) <- cbind("so_af_gdp_diff","so_af_claims_diff","so_af_cpi_diff","so_af_milt_diff")

```


## Train, test split

```{r, echo=FALSE}

split_so_af <- ts_split(ts.obj = so_af_vec, sample.out = 4)

train_so_af <- split_so_af$train
test_so_af <- split_so_af$test

length(train_so_af)
length(test_so_af)

```


# VAR - Vector AutoRegressive model


## Find optimal lag length

```{r, echo=FALSE}

lagselect <- VARselect(so_af_vec, lag.max = 10, type = "const")
lagselect$selection

```


## Build VAR model

```{r, echo=FALSE}

so_af_model <- VAR(so_af_vec, p = 4, type = "const", season = NULL, exog = NULL) 
#summary(so_af_model)

```


## Stargazer Output

```{r, echo=FALSE}

stargazer(so_af_model[["varresult"]], type="text")

```


## Check for Serial Correlation

```{r}

so_af_serial <- serial.test(so_af_model, type = "PT.asymptotic")
so_af_serial

```


## Test for Heteroscedasticity

```{r}

so_af_arch <- vars::arch.test(so_af_model, lags.multi = 15, multivariate.only = TRUE)
so_af_arch

```


## Test for Normality of Residuals

```{r}

so_af_norm <- normality.test(so_af_model, multivariate.only = TRUE)
so_af_norm
#plot(so_af_model)

```


## Check for Stability of model

```{r}

so_af_stability <- stability(so_af_model, type = "OLS-CUSUM")
plot(so_af_stability)

```


## Check for Granger Causality

```{r}

granger_gdp <- causality(so_af_model, cause = "so_af_gdp_diff")
granger_gdp

granger_claims <- causality(so_af_model, cause = "so_af_claims_diff")
granger_claims

granger_cpi <- causality(so_af_model, cause = "so_af_cpi_diff")
granger_cpi

granger_milt <- causality(so_af_model, cause = "so_af_milt_diff")
granger_milt

```



## Create Impulse Response Functions

```{r}

# Black line outside of confidence interval means the response is significant!
so_af_claims_irf <- irf(so_af_model, impulse = "so_af_claims_diff", response = "so_af_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(so_af_claims_irf)

so_af_cpi_irf <- irf(so_af_model, impulse = "so_af_cpi_diff", response = "so_af_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(so_af_cpi_irf)

so_af_milt_irf <- irf(so_af_model, impulse = "so_af_milt_diff", response = "so_af_gdp_diff", n.ahead = 4, boot = TRUE, run = 200, ci = 0.9, ortho = FALSE)
plot(so_af_milt_irf)

```



## Review Variance Decomposition

```{r}

vd1 <- fevd(so_af_model, n.ahead = 4)

plot(vd1)

```



## Generate Predictions

```{r}

forecast <- predict(so_af_model, n.ahead = 4, ci = 0.95)
fanchart(forecast, names = "so_af_gdp_diff", colors = c("seagreen","lightgreen"), main = "Forecasts", xlab = "Year", ylab = "GDP Growth %")

```

# View Actuals vs Predicted

```{r}

index_pred <- append(index(train_so_af[,1]), 2017:2019)

all_actuals <- append(train_so_af[,1], test_so_af[1:3,1])

all_preds <- append(train_so_af[,1], forecast$fcst$so_af_gdp_diff[,1])

all_actuals_dediff <- diffinv(all_actuals, xi = so_af_gdp[1])

all_preds_dediff <- diffinv(all_preds, xi = so_af_gdp[1])
all_preds_dediff <- all_preds_dediff[-length(all_preds_dediff)]

all_actuals_dediff
all_preds_dediff

length(index_pred)
length(all_actuals_dediff)
length(all_preds_dediff[-1])

plot(index_pred,
     all_preds_dediff[-1],
     type = "l",
     ylab="Actuals and Predicted",
     xlab="Year",
     col = "red")
lines(index_pred,
      all_actuals_dediff[-1],
      type = "l",
      col = "green")
legend("topleft",
c("Predicted","Actuals"),
fill=c("red","green"))

```

## VAR - Evaluation Metrics

```{r}

# Root Mean Squared Error
print("Root Mean Squared Error")
rmse(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Error
print("Mean Absolute Error")
mae(all_actuals_dediff[37:39], all_preds_dediff[37:39])


# Mean Absolute Percent Error
print("Mean Absolute Percent Error")
mape(all_actuals_dediff[37:39], all_preds_dediff[37:39])

```











